---
title: "Figures for \"Perturbed progenitor cell behavior and YAP mediated expansion drives EML1-linked cerebral cortical malformation in a human forebrain organoid model.\""
author: "Anne Hoffrichter"
date: "2021/07/12"
output: 
  bookdown::html_document2:
    code_folding: show
    fig_caption: true
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts=list(width.cutoff=80),tidy=TRUE, fig.asp=0.5, fig.width=8, warning = FALSE, dev = "pdf")
```

```{r loadLibraries, message=FALSE, warning=FALSE}
library(Seurat)
library(tidyverse)
library(patchwork)
library(org.Hs.eg.db)
library(GO.db)
options(stringsAsFactors = FALSE)
set.seed(42)
```


```{r}
data <- readRDS("../data/integrated_data.rds")
DefaultAssay(data) <- "RNA"
```

```{r}
my_levels <- c("RG1", "RG to bRG", "RG2", "IP", "YN", "MLC")
Idents(data) <- factor(Idents(data), levels = my_levels)
```

# Figure 2 - ScRNA-seq reveals cellular identity of ectopic cells in EML1-KO cerebral organoids
## a) Dimensional reduction plot (UMAP)

scRNA-seq of 2,335 control and 3,358 EML1-KO cells (3 pooled independent batches, 3 organoids each). Cells shown in UMAP plot and colored by annotated cell type.

```{r, fig.asp=1}
a <- subset(data, KO=="CTRL") %>% DimPlot(pt.size = 0.2)+
  ggtitle("Control")+
  NoLegend()+ 
  theme(plot.title = element_text(face="plain", size=20, hjust=0.5))
b <- subset(data, KO=="KO") %>% DimPlot(pt.size = 0.2)+
  ggtitle(substitute(paste(italic("EML1"),"-KO")))+
  theme(plot.title = element_text(face="plain", size=20, hjust=0.5))
```

## b) DotPlot - marker genes for specifying cell types

Dotplot of known marker gene expression in control-derived cells. Scaled average normalized expression for the respective gene in each cell type is indicated by a continuous color scale. Number of cells in the respective cluster expressing the genes is represented by the size of the dots. 

```{r}
genes <- c("CDH2", "C1orf61", "FABP7", "HES1", "VIM", "SOX2", "NES", "OTX2", "PAX6", "EMX2", 
  "SFRP1", "ANXA5", "PTPRZ1", "GSN", "FABP5",
  "LUM", "DCN", "COL5A1",
  "NEUROD1", "EOMES",  "ELAVL4", "NHLH1", 
  "MEIS2", "POU3F2", "TBR1", "MAPT", "STMN1", "DCX", "SNAP25" 
           )
```

```{r}
my_levels <- c("MLC", "YN", "IP", "RG2", "RG to bRG", "RG1")
Idents(data) <- factor(Idents(data), levels = my_levels)


c <-subset(data, KO=="CTRL") %>% DotPlot(features = genes, cols=c("lightgrey", "#DA5B51"))+ 
  theme_classic(base_size=13)+
  RotatedAxis()+
  theme(plot.title = element_text(hjust = 0.5, face="plain"), axis.title = element_blank(), legend.text = element_text(size = 8), legend.title=element_text(size = 8), legend.key.size = unit(0.3, "cm"))+
  guides(color = guide_colorbar(title = 'Average\nExpression'),size=guide_legend(title="Percent\nExpressed"))+
  geom_vline(xintercept = c(10.5, 15.5, 18.5, 22.5), linetype="dotted", size=0.5)
```

## c) percent cells per cell type Ctrl vs. KO without MLC

Comparison of cell-type composition between control and EML1-KO. Numbers indicate percentage of total cells belonging to the respective cell-type. 

```{r}
my_levels <- rev(my_levels)
Idents(data) <- factor(Idents(data), levels = my_levels)
data_subset <- subset(data, cell_type != "MLC")

Ctrl <- subset(data_subset, KO=="CTRL")
KO <- subset(data_subset, KO=="KO")
Idents(Ctrl) <- factor(Idents(Ctrl), levels = my_levels)
Idents(KO) <- factor(Idents(KO), levels = my_levels)
tCtrl <- table(Ctrl$cell_type)
tKO <- table(KO$cell_type)
df <-data.frame(tCtrl, tKO)
df <- df[c(1:4,6),c(1:2,4)]

colnames(df) <- c("cell_type", "Control", "KO")
pct_Ctrl <- round(df$Control/sum(df$Control)*100, digits = 1)
pct_KO <- round(df$KO/sum(df$KO)*100, digits = 1)
pct <- data.frame(pct_Ctrl, pct_KO, row.names = df$cell_type)
pct2 <- data.frame(KO="Control", pct=pct$pct_Ctrl, cell_type=rownames(pct))
pct3 <- data.frame(KO="KO", pct=pct$pct_KO, cell_type=rownames(pct))
pct4 <- rbind(pct2,pct3)
pct4$cell_type <- ordered(pct4$cell_type, levels=my_levels[1:5])
```

```{r, fig.asp=1, fig.width=6}
d <- ggplot(data=pct4, aes(x=cell_type, y=pct, fill=KO)) + 
  geom_bar(stat="identity", position=position_dodge2(reverse=T))+
  coord_flip()+
  geom_text(aes(label=paste0(pct, " %")),hjust=- 0.1, color="black", position = position_dodge2(0.9, reverse=T), size=3)+
  theme_minimal(base_size = 13)+ 
  scale_x_discrete(limits=rev(levels(pct4$cell_type)))+ 
  theme(axis.text = element_text(colour = "black", face="plain")) +
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  ylim(0,65)+
  labs(x = "", y="", fill=c("Control", substitute(paste(italic("EML1"),"-KO"))))+ 
  labs(fill = "")
```

## Figure 2 panel a & b & c

```{r, fig.width=14, fig.asp=0.65}
a+c+b+d+
  plot_layout(widths = c(1, 2))
```

## d) & i) MEIS2, COL1A2, COL3A1, LUM violin plots without MLCs

Normalized expression values of MEIS2 and COL1A2 per cell-type. 

Normalized expression values of COL3A1 and LUM per cell-type.

```{r}
my_levels <- c("RG1", "RG to bRG", "RG2", "IP", "YN")
genes <- c("MEIS2", "COL1A2", "COL3A1", "LUM")
Idents(data_subset) <- "cell_type"

sub_list <- map(my_levels, ~subset(data_subset, cell_type==.x))
for(i in seq_along(sub_list)){
  Idents(sub_list[[i]]) <- "KO"
}

m <- map(sub_list, ~FindMarkers(object=.x,ident.1 = "KO", ident.2 = "CTRL", features = genes, logfc.threshold = 0, min.pct=0))
names(m) <- my_levels
```

```{r}
m <- map2(my_levels, m, ~cbind(.y, cell_type=.x))
m <- map(m, ~cbind(.x, gene=rownames(.x)))
m_df <- rbind(m[[1]], m[[2]], m[[3]],m[[4]], m[[5]])
for(i in seq_along(m_df$p_val_adj)){
  if(m_df$p_val_adj[i]<0.001){
    m_df$label[i] <- "***"
  }else if (m_df$p_val_adj[i]<0.01){
    m_df$label[i] <- "**"
  }else if (m_df$p_val_adj[i]<0.05){
    m_df$label[i] <- "*"
  }else {
    m_df$label[i] <- "ns"
  }
}
```


```{r}
# Stats
m_df
```


```{r}
Idents(data_subset) <- "cell_type"
Idents(data_subset) <- factor(Idents(data_subset), levels = my_levels)
```

```{r, fig.asp=0.25, fig.width=12}

p <- vector("list", 4)

p[[1]] <- VlnPlot(data_subset, features = genes[1], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[1],])+max(GetAssayData(data_subset, slot="data")[genes[1],])*0.1, label=c("ns", "***", "***", "ns", "ns"), size=c(4,7,7,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[1],])+max(GetAssayData(data_subset, slot="data")[genes[1],])*0.15)


p[[2]] <- VlnPlot(data_subset, features = genes[2], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[2],])+max(GetAssayData(data_subset, slot="data")[genes[2],])*0.1, label=c("***", "***", "ns", "ns", "ns"), size=c(7,7,4,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[2],])+max(GetAssayData(data_subset, slot="data")[genes[2],])*0.15)


p[[3]]<- VlnPlot(data_subset, features = genes[3], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[3],])+max(GetAssayData(data_subset, slot="data")[genes[3],])*0.1, label=c("***", "***", "*", "ns", "ns"), size=c(7,7,7,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[3],])+max(GetAssayData(data_subset, slot="data")[genes[3],])*0.15)


p[[4]] <- VlnPlot(data_subset, features = genes[4], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[4],])+max(GetAssayData(data_subset, slot="data")[genes[4],])*0.1, label=c("***", "***", "ns", "ns", "ns"), size=c(7,7,4,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[4],])+max(GetAssayData(data_subset, slot="data")[genes[4],])*0.15)

  

(p3 <- wrap_plots(p, ncol=4)+
    plot_layout(guides = "collect"))

```

## j) GO violin plots RG to bRG

Gene ontology (GO) term analysis of RG to bRG cells in control and EML1-KO shows percentage of counts belonging to the set of genes associated with the respective GO term.

```{r}
add_GO_perc <- function(data, GO_term){
  go_id = GOID(GOTERM[Term(GOTERM) == GO_term])
  allegs <- get(go_id, org.Hs.egGO2ALLEGS)
  genes <- unlist(mget(allegs, org.Hs.egSYMBOL))
  genes <- unique(genes)        
  genes <- genes[genes %in% rownames(data)]
  PercentageFeatureSet(object=data, features=genes, col.name = GO_term)
}  
```

```{r}
data_subset <- ScaleData(data_subset)
DefaultAssay(data_subset) <- "RNA"
RGtobRG <- subset(data_subset, cell_type=="RG to bRG")
RGtobRG <- ScaleData(RGtobRG)
DefaultAssay(RGtobRG) <- "RNA"
Idents(RGtobRG) <- "KO"
```

```{r, fig.asp=0.22, fig.width=16}
GO_terms <- c( "positive regulation of neural precursor cell proliferation","extracellular matrix assembly", "cytoplasmic microtubule organization", "establishment of mitotic spindle orientation", "cilium assembly")

for(i in seq_along(GO_terms)){
  RGtobRG <- add_GO_perc(RGtobRG, GO_terms[i])
}

titles <- c("Positive regulation of neural \n precursor cell proliferation", "Extracellular matrix \n assembly", "Cytoplasmic microtubule \n organization", "Establishment of mitotic \n spindle orientation", "Cilium assembly")

GO_terms_no_space <- str_replace_all(GO_terms, " ", ".")

plots3 <- map2(GO_terms_no_space, titles, ~VlnPlot(RGtobRG, features=.x, split.by = "KO", pt.size=0.1)+ ggtitle(.y))
```

```{r}
# Stats
wil_t <- list()
for(i in GO_terms_no_space){
  df <- data.frame(GO=RGtobRG[[i]], KO=RGtobRG$KO)
  wil_t[[i]] <- wilcox.test(x=filter(df, KO=="CTRL")[,1], y=filter(df, KO=="KO")[,1])
}
wil_t
```

```{r}
plots3[[1]] <- plots3[[1]]+
  annotate("text", x=1.5, y=max(RGtobRG[[GO_terms_no_space[1]]])+max(RGtobRG[[GO_terms_no_space[1]]])*0.1, label="***", size=10)+
  ylim(min(RGtobRG[[GO_terms_no_space[1]]])-min(RGtobRG[[GO_terms_no_space[1]]])*0.15,max(RGtobRG[[GO_terms_no_space[1]]])+max(RGtobRG[[GO_terms_no_space[1]]])*0.15)

plots3[[2]] <- plots3[[2]]+
  annotate("text", x=1.5, y=max(RGtobRG[[GO_terms_no_space[2]]])+max(RGtobRG[[GO_terms_no_space[2]]])*0.1, label="***", size=10)+
  ylim(min(RGtobRG[[GO_terms_no_space[2]]])-min(RGtobRG[[GO_terms_no_space[2]]])*0.15,max(RGtobRG[[GO_terms_no_space[2]]])+max(RGtobRG[[GO_terms_no_space[2]]])*0.15)

plots3[[3]] <- plots3[[3]]+
  annotate("text", x=1.5, y=max(RGtobRG[[GO_terms_no_space[3]]])+max(RGtobRG[[GO_terms_no_space[3]]])*0.1, label="***", size=10)+
  ylim(min(RGtobRG[[GO_terms_no_space[3]]])-min(RGtobRG[[GO_terms_no_space[3]]])*0.15,max(RGtobRG[[GO_terms_no_space[3]]])+max(RGtobRG[[GO_terms_no_space[3]]])*0.15)

plots3[[4]] <- plots3[[4]]+
  annotate("text", x=1.5, y=max(RGtobRG[[GO_terms_no_space[4]]])+max(RGtobRG[[GO_terms_no_space[4]]])*0.1, label="***", size=10)+
  ylim(min(RGtobRG[[GO_terms_no_space[4]]])-min(RGtobRG[[GO_terms_no_space[4]]])*0.15,max(RGtobRG[[GO_terms_no_space[4]]])+max(RGtobRG[[GO_terms_no_space[4]]])*0.15)

plots3[[5]] <- plots3[[5]]+
  annotate("text", x=1.5, y=max(RGtobRG[[GO_terms_no_space[5]]])+max(RGtobRG[[GO_terms_no_space[5]]])*0.1, label="***", size=10)+
  ylim(min(RGtobRG[[GO_terms_no_space[5]]])-min(RGtobRG[[GO_terms_no_space[5]]])*0.15,max(RGtobRG[[GO_terms_no_space[5]]])+max(RGtobRG[[GO_terms_no_space[5]]])*0.15)
```



```{r, fig.asp=0.28, fig.width=16}
wrap_plots(plots3, guides="collect", nrow=1)& 
  theme(axis.title.x=element_blank(), axis.text.y=element_text(size=18), plot.title = element_text(size=14), legend.text = element_text(size=14), axis.title.y=element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank())&
  ylab("Proportion of feature set")&
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))
```

# Figure 4
## b) Violin Plots YAP, CCND3, CDK6, CYR61 only RG to bRG cells

Normalized expression of YAP1, TEAD2, CCND3, and CYR61 in RG to bRG cells. Separate violins show expression in control and EML1-KO. Asterisks indicate Bonferroni corrected p-values. Percentages indicate amount of cells in each group expressing the respective gene.

```{r}
Idents(RGtobRG) <- "KO"
genes <- c("YAP1", "TEAD2", "CCND3", "CYR61")
```

```{r}
m <- FindMarkers(RGtobRG, ident.1 = "KO", ident.2 = "CTRL", features = genes, logfc.threshold = 0)

for(i in seq_along(m$p_val_adj)){
  if(m$p_val_adj[i]<0.001){
    m$label[i] <- "***"
  }else if (m$p_val_adj[i]<0.01){
    m$label[i] <- "**"
  }else if (m$p_val_adj[i]<0.05){
    m$label[i] <- "*"
  }else {
    m$label[i] <- "ns"
  }
}
```

```{r}
# Stats
m
```



```{r, fig.asp=0.6, fig.width=6}
p <- vector("list", 4)

p[[1]] <- VlnPlot(RGtobRG, features = genes[1], split.by = "KO", pt.size = 0)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=1.5, y=max(GetAssayData(RGtobRG, slot="data")[genes[1],])+max(GetAssayData(RGtobRG, slot="data")[genes[1],])*0.1, label="*", size=10)+
  ylim(-0.1,max(GetAssayData(RGtobRG, slot="data")[genes[1],])+max(GetAssayData(RGtobRG, slot="data")[genes[1],])*0.15)+
  annotate("text", x=1, y=max(GetAssayData(RGtobRG, slot="data")[genes[1],])+max(GetAssayData(RGtobRG, slot="data")[genes[1],])*0.05, label=paste0(m[genes[1], "pct.2"]*100, "%"), size= 5)+
  annotate("text", x=2, y=max(GetAssayData(RGtobRG, slot="data")[genes[1],])+max(GetAssayData(RGtobRG, slot="data")[genes[1],])*0.05, label=paste0(m[genes[1], "pct.1"]*100, "%"), size= 5)

p[[2]] <- VlnPlot(RGtobRG, features = genes[2], split.by = "KO", pt.size = 0)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"), axis.title.y = element_blank())+
  annotate("text", x=1.5, y=max(GetAssayData(RGtobRG, slot="data")[genes[2],])+max(GetAssayData(RGtobRG, slot="data")[genes[2],])*0.1, label="***", size=10)+
  ylim(-0.1,max(GetAssayData(RGtobRG, slot="data")[genes[2],])+max(GetAssayData(RGtobRG, slot="data")[genes[2],])*0.15)+
  annotate("text", x=1, y=max(GetAssayData(RGtobRG, slot="data")[genes[2],])+max(GetAssayData(RGtobRG, slot="data")[genes[2],])*0.05, label=paste0(m[genes[2], "pct.2"]*100, "%"), size= 5)+
  annotate("text", x=2, y=max(GetAssayData(RGtobRG, slot="data")[genes[2],])+max(GetAssayData(RGtobRG, slot="data")[genes[2],])*0.05, label=paste0(m[genes[2], "pct.1"]*100, "%"), size= 5)

p[[3]]<- VlnPlot(RGtobRG, features = genes[3], split.by = "KO", pt.size = 0)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"), axis.title.y = element_blank())+
  ylim(-0.1,max(GetAssayData(RGtobRG, slot="data")[genes[3],])+max(GetAssayData(RGtobRG, slot="data")[genes[3],])*0.15)+
  annotate("text", x=1.5, y=max(GetAssayData(RGtobRG, slot="data")[genes[3],])+max(GetAssayData(RGtobRG, slot="data")[genes[3],])*0.13, label="***", size=10)+
  
  annotate("text", x=1, y=max(GetAssayData(RGtobRG, slot="data")[genes[3],])+max(GetAssayData(RGtobRG, slot="data")[genes[3],])*0.05, label=paste0(m[genes[3], "pct.2"]*100, "%"), size= 5)+
  annotate("text", x=2, y=max(GetAssayData(RGtobRG, slot="data")[genes[3],])+max(GetAssayData(RGtobRG, slot="data")[genes[3],])*0.05, label=paste0(m[genes[3], "pct.1"]*100, "%"), size= 5)

p[[4]] <- VlnPlot(RGtobRG, features = genes[4], split.by = "KO", pt.size = 0)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"), axis.title.y = element_blank())+
  annotate("text", x=1.5, y=max(GetAssayData(RGtobRG, slot="data")[genes[4],])+max(GetAssayData(RGtobRG, slot="data")[genes[4],])*0.1, label="**", size=10)+
  ylim(-0.1,max(GetAssayData(RGtobRG, slot="data")[genes[4],])+max(GetAssayData(RGtobRG, slot="data")[genes[4],])*0.15)+
  annotate("text", x=1, y=max(GetAssayData(RGtobRG, slot="data")[genes[4],])+max(GetAssayData(RGtobRG, slot="data")[genes[4],])*0.05, label=paste0(m[genes[4], "pct.2"]*100, "%"), size= 5)+
  annotate("text", x=2, y=max(GetAssayData(RGtobRG, slot="data")[genes[4],])+max(GetAssayData(RGtobRG, slot="data")[genes[4],])*0.05, label=paste0(m[genes[4], "pct.1"]*100, "%"), size= 5)
```

```{r, fig.asp=0.6, fig.width=9}
(p3 <- wrap_plots(p, ncol=4)+
    plot_layout(guides = "collect"))  & theme(legend.position = 'bottom')
```


# Figure S2
## b) Violin Plot EML1 Ctrl 

Normalized expression of EML1 in day 33 ± 2 control-derived cerebral organoid cells per cell type investigated by scRNAseq.

```{r}
Idents(Ctrl) <- "cell_type"
genes <- c("EML1")
my_levels <- c("RG1", "RG to bRG", "RG2", "IP", "YN")
Idents(Ctrl) <- factor(Idents(Ctrl), levels = my_levels)
```

```{r, fig.asp=0.5, fig.width=6}
VlnPlot(Ctrl, features = genes[1], pt.size = 0)+
  scale_fill_manual(values=rep("#DA5B51", times=5))+
  theme(axis.title.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  NoLegend()
```


## g) Heatmap - RG to bRG (without MLC)

Heat map of z-scaled gene expression of extracellular matrix (ECM) and basal radial glial (bRG) genes in the RG to bRG cell cluster derived from either control or EML1-KO organoids.

```{r}
genes <- c("COL1A2", "COL3A1", "LUM", 
            "COL1A1", "COL5A2", "COL6A2", "COL6A3", "COL13A1", "ANXA6", "APOE", "BGN", "CCDC80", "CD248", "CXCL12", "DCN", "EFEMP2", "EMILIN1", "FBLN1", "FBLN2", "FN1", "GPC3", "LOXL1", "MDK", "MFAP2", "MFAP4", "MMP11", "MMP2", "PCOLCE", "POSTN", "SERPINF1", "TGFB1", "TIMP1", "TRIL", "S100A4", "SERPINB6", "SERPINH1",
           "MOXD1","FBN2", "PLOD1", "GPC6", "NDRG2", "PTN", "ETV5", "LITAF", "TFPI", "LRRC17", "ITM2C", "OAT", "ANXA2", "GNG11", "RHOC", "IFI6", "LIMA1", "TPM2",  "LGALS1", "MFGE8", "SFRP1")
```

```{r, fig.asp=0.8, dev = "png"}
DoHeatmap(RGtobRG, features = genes, group.colors = c("#DA5B51", "#00A5AA"), label = FALSE, draw.lines = FALSE)+
  scale_fill_viridis_c(option="viridis")+
  theme(axis.text.y = element_text(face="bold.italic"))+
  geom_hline(yintercept = c(21.5), color="white")+
  annotate("text", x = 220, y=60.5, label = "Control")+
  annotate("text", x= 1114, y = 60.5, label = substitute(paste(italic("EML1"),"-KO")))
```


# Figure S4
## g) ViolinPlots YAP1, CCND3, CDK6, CYR61 all cell types without MLCs

Normalized expression of YAP1, CCND3, CDK6 and CYR61 in all cell types. Split violins show expression in control and EML1-KO.

```{r}
my_levels <- c("RG1", "RG to bRG", "RG2", "IP", "YN")
genes <- c("YAP1", "TEAD2", "CCND3", "CYR61")
Idents(data_subset) <- "cell_type"
Idents(data_subset) <- factor(Idents(data_subset), levels = my_levels)
sub_list <- map(my_levels, ~subset(data_subset, cell_type==.x))
for(i in seq_along(sub_list)){
  Idents(sub_list[[i]]) <- "KO"
}

m <- map(sub_list, ~FindMarkers(object=.x,ident.1 = "KO", ident.2 = "CTRL", features = genes, logfc.threshold = 0, min.pct=0))
names(m) <- my_levels
```

```{r}
m <- map2(my_levels, m, ~cbind(.y, cell_type=.x))
m <- map(m, ~cbind(.x, gene=rownames(.x)))
m_df <- rbind(m[[1]], m[[2]], m[[3]],m[[4]], m[[5]])
for(i in seq_along(m_df$p_val_adj)){
  if(m_df$p_val_adj[i]<0.001){
    m_df$label[i] <- "***"
  }else if (m_df$p_val_adj[i]<0.01){
    m_df$label[i] <- "**"
  }else if (m_df$p_val_adj[i]<0.05){
    m_df$label[i] <- "*"
  }else {
    m_df$label[i] <- "ns"
  }
}
```


```{r}
# Stats
m_df
```


```{r, fig.asp=0.75, fig.width=6}
p <- vector("list", 4)

p[[1]] <- VlnPlot(data_subset, features = genes[1], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[1],])+max(GetAssayData(data_subset, slot="data")[genes[1],])*0.1, label=c("ns", "*", "ns", "ns", "ns"), size=c(4,7,4,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[1],])+max(GetAssayData(data_subset, slot="data")[genes[1],])*0.15)

p[[2]] <- VlnPlot(data_subset, features = genes[2], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[2],])+max(GetAssayData(data_subset, slot="data")[genes[2],])*0.1, label=c("ns", "***", "ns", "ns", "ns"), size=c(4,7,4,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[2],])+max(GetAssayData(data_subset, slot="data")[genes[2],])*0.15)

p[[3]]<- VlnPlot(data_subset, features = genes[3], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[3],])+max(GetAssayData(data_subset, slot="data")[genes[3],])*0.1, label=c("ns", "***", "ns", "ns", "ns"), size=c(4,7,4,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[3],])+max(GetAssayData(data_subset, slot="data")[genes[3],])*0.15)

p[[4]] <- VlnPlot(data_subset, features = genes[4], split.by = "KO", pt.size = 0, split.plot = TRUE)+
  scale_fill_manual(values=c("#DA5B51", "#00A5AA"), labels=c("Control", substitute(paste(italic("EML1"),"-KO"))))+
  theme(axis.title.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(face="bold.italic"))+
  annotate("text", x=c(1:5), y=max(GetAssayData(data_subset, slot="data")[genes[4],])+max(GetAssayData(data_subset, slot="data")[genes[4],])*0.1, label=c("ns", "**", "ns", "ns", "ns"), size=c(4,7,4,4,4))+
  ylim(-0.1,max(GetAssayData(data_subset, slot="data")[genes[4],])+max(GetAssayData(data_subset, slot="data")[genes[4],])*0.15)
```

```{r, fig.asp=0.8, fig.width=9}
(p3 <- wrap_plots(p, ncol=2)+
    plot_layout(guides = "collect"))  & theme(legend.position = 'bottom')
```









```{r}
sessionInfo()
```

